---
title: "prepare_results"
author: "ab"
date: "`r Sys.Date()`"
output: html_document
---

Notebook with a code to generate plots for electrophysiology assay


```{r Source libraries}
source("./F_prepare.R")
options(show_col_types = F)
```

```{r Prepare results dir}
if (!dir.exists("./results")) {
  dir.create("./results")
}
prfx <- ""  # can add a name to you run or just leave empty
```

First, we want to load in the data and clean it bit for the analysis
```{r 1. Load data}
t_spikes <- read_tsv("./processed/t_spikes.tsv", show_col_types = F)
t_bursts <- read_tsv("./processed/t_bursts.tsv", show_col_types = F)
t_ns <- read_tsv("./processed/t_ns.tsv", show_col_types = F)
```


# ------------------------------------------------------------------------------
## 1. Spike rate analysis

### Manual filtering
```{r}
## filter inactive wells cells in D3 and D2 die at day 21, only residual activity at day 25
t_spikes_filt <- t_spikes %>% 
  filter(!(Well %in% c("D3", "D2") & DIV == "25"),
         Treatment %in% c("Scrambled", "ShYTFD"), 
         !Well %in% c("D5"))

## summarise per DIV for line plotting
t_spikes_filt_sum <- t_spikes_filt %>%  
  group_by(Treatment, DIV) %>% 
  summarise(MFR_br.avg = mean(MFR, na.rm = T),
            MFR_br.sd = sd(MFR, na.rm = T),
            .groups = "drop") %>% 
  mutate(DIV = as.numeric(DIV))
```

### Plotting
```{r}
## quickie to make outliers coloured
update_geom_defaults("point", list(colour = NULL))
ggplot(data = t_spikes_filt %>% 
         mutate(DIV = factor(DIV, levels = c(7, 11, 14, 18, 21))), 
       aes(x = DIV, y = MFR)) + 
  geom_boxplot(aes(fill = Treatment), size = 0.25, outlier.shape = 21) + 
  #scale_fill_manual(values = alpha(pal_mut.msa[c(1,3)], 0.67)) + 
  theme(aspect.ratio = 1.25) + 
  labs(y = "MFR (Hz / nAE)") -> p.mfr_boxes
## revert to default options for durther plotting
update_geom_defaults("point", list(colour = "black"))
```

```{r}
ggplot(data = t.t_spikes_filt_sum, 
         aes(x = DIV, y = MFR_br.avg, color = Treatment)) + 
    geom_point(aes(shape = Treatment), 
               size = 2.5, 
               position = position_dodge(width = 0.35)) + 
    geom_errorbar(aes(ymax = MFR_br.avg + MFR_br.sd,
                      ymin = MFR_br.avg - MFR_br.sd),
                  size = 0.5,
                  width = 0.25,
                  position = position_dodge(width = 0.35)) + 
  geom_line(position = position_dodge(width = 0.35), size = 1) + 
  #geom_hline(yintercept = 0) +
  scale_color_manual(values = alpha(pal_mut.msa[c(1,3)], 0.67)) + 
  scale_x_continuous(breaks = seq(0, 25, by = 1)) +
  theme(aspect.ratio = 0.5 ,
        legend.position = c(0.2, 0.75)) + 
  coord_cartesian(ylim = c(0, 50)) +
  labs(y = "MFR, HZ / nAE", x = "DIV") -> p.mfr_lines
```

```{r fig.width=4.5, fig.height=3}
ggsave(plot = p.mfr_lines, filename = paste0("./results/", prfx, "MFR.lines.pdf"),
       width = 4.5, height = 3)
```


### Permutation test ----
To test unambigously test for the differences between case/control we perform a MWU-test 5000 times
on the same data with randomly permuted labels
```{r}
t_spikes_filt %>%
  filter(DIV %in% c("14", "18", "21", "25")) %>% 
  wilcox.test(formula = MFR ~ Treatment, data = .) -> spks.test_mwu

t_spikes_filt %>%
  filter(DIV %in% c("14", "18", "21", "25")) %>% 
  perm_test(tab = ., col_to_perm = "Treatment", col_to_test = "MFR", n_perm = 5000) -> t.perm

t.perm <- t.perm %>% 
  mutate(m.lg10_pval = -log10(p.value))

# ggplot(data = t.perm, 
#        aes(x = m.lg10_pval, y = ..density..)) + 
#   geom_histogram(bins = 50, fill = NA, size = 0.5, color = "black") + 
#   geom_density(color = "red")
diff(ecdf(t.perm$m.lg10_pval)(c(-log10(spks.test_mwu$p.value), Inf)))
```

## Summary of activity on a plate
```{r}
t_spikes_filt %>% 
  #mutate(DIV = as.character(DIV)) %>% 
  mutate(DIV = factor(DIV, levels = c(7, 11, 14, 18, 21, 25))) %>% 
  group_by(Treatment, DIV) %>% 
  summarise(n_wells = n(),
            n_ae = sum(nae),
            .groups = "drop") -> t.yth_summary
write_tsv(t_spikes_filt_summary, paste0("./results/", prfx, "nAE.summary.tsv"))
```





# ------------------------------------------------------------------------------
## 2. Burst rate analysis
### Manual filtering (same as before)
```{r}
t_bursts_filt <- t_bursts %>% 
  filter(!(Well %in% c("D3", "D2") & DIV == "25"),
         Treatment %in% c("Scrambled", "ShYTFD"), 
         !Well %in% c("D5")) %>% 
  mutate(DIV = factor(DIV, levels = c(7, 11, 14, 18, 21, 25)))
```
```{r}
## quickie to make outliers coloured
update_geom_defaults("point", list(colour = NULL))
ggplot(data = t_bursts_filt, 
       aes(x = DIV, y = brst.size_avg)) + 
  geom_boxplot(aes(fill = Treatment), outlier.shape = 21) + 
  scale_fill_manual(values = alpha(pal_mut.msa[c(1,3)], 0.67)) + 
  scale_y_log10() +
  theme(aspect.ratio = 1.5) + 
  labs(y = "Mean burst size, spikes") -> p.burstsize

## revert to default options for durther plotting
update_geom_defaults("point", list(colour = "black"))
```

```{r fig.width=3.5, fig.height=3}
ggsave(plot = p.burstsize, 
       filename = paste0("./results/", prfx, "MBS.boxes.pdf"),
       width = 3.5, height = 3)
```

### Permutation test ----
To test for the differences between case/control we perform a MWU-test 5000 times
on the same data with randomly permuted labels
```{r}
t_bursts_filt %>%
  filter(DIV %in% c("14", "18", "21", "25")) %>% 
  wilcox.test(formula = brst.size_avg ~ Treatment, data = .) -> brsts.test_mwu

t_bursts_filt %>%
  filter(DIV %in% c("14", "18", "21", "25")) %>% 
  perm_test(tab = ., col_to_perm = "Treatment", col_to_test = "brst.size_avg", n_perm = 5000) -> t.perm_brsts

t.perm_brsts <- t.perm_brsts %>% 
  mutate(m.lg10_pval = -log10(p.value))

# ggplot(data = t.perm, 
#        aes(x = m.lg10_pval, y = ..density..)) + 
#   geom_histogram(bins = 50, fill = NA, size = 0.5, color = "black") + 
#   geom_density(color = "red")
diff(ecdf(t.perm$m.lg10_pval)(c(-log10(brsts.test_mwu$p.value), Inf))) # <----- here we differentiate the p.dist and get the p-value for perm.test
```


# ------------------------------------------------------------------------------
## 3. Snap shot of bursting from selected region

### Read in the firing data
```{r}
l.div <- read_rds("./processed/l_assembled.RDS")

## I will look specifically at DIV18 as it has the largest differences in spike features
div18 <- l.div$`18`
## Only keep wells that are relevant to the analysis
div18 <- div18[t.yth[t.yth$DIV == "18", ]$Well]

purrr::map2_dfr(div18_mhz,
            names(div18_mhz),
            function(fr_list, well) {
              fr_list$firing_rates$frate %>% 
                as_tibble(., rownames = "bin") %>% 
                pivot_longer(., !contains("bin"), names_to = "Electrode", values_to = "spikes") %>% 
                mutate(Well = well) %>% 
                dplyr::select(Well, Electrode, bin, spikes)
            }) %>% 
  inner_join(t.meta, ., by = "Well") -> t.raster_mhz
```

```{r}
ggplot(data = t.raster_mhz %>% 
         mutate(bin = factor(bin, levels = 0:9999),
                Electrode = paste0(Well, "_", Electrode)), 
       aes(y = Electrode, x = bin)) + 
  geom_raster(aes(fill = spikes)) +
  #geom_vline(xintercept = c(200, 280)) +
  facet_grid(Treatment + Well ~ ., scales = "free") + 
  # facet_grid(Treatment + Well ~ ., 
  #            scales = "free_y", 
  #            nrow = 2, 
  #            strip.position = "left", 
  #            labeller = as_labeller(c(Scrambled = "Scrambled\nElectrodes 1 -> 51", 
  #                                     ShYTFD = "shYthdf123\nElectrodes 1 -> 43")))  +
  scale_fill_gradient2(low = "white", high = "black") + 
  scale_x_discrete(breaks = seq(0, 10000, by = 1000), labels = seq(0, 100, by = 10)) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(), 
        axis.line.y = element_blank(),
        panel.border = element_rect(size = 0.25, fill = NA)) +
  labs(y = NULL, x = "Recording time, s") +
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        legend.position = "none") -> p.raster_zoomed
```
```{r fig.width=8, fig.height=6}
ggsave(plot = p.raster_zoomed, 
       filename = paste0("./results/", prfx, "Raster.zoomed.pdf"),
       width = 8, height = 6)
```


### Zoom in to individual electrodes
Pick representative electrodes for final plottings
Extremes are D1 and C1, so we pick D2 and B4
```{r, fig.height=2, fig.width=5}
t.raster_mhz_sniped <- t.raster_mhz %>% 
  filter(Well %in% c("D2", "B4")) %>% 
  mutate(bin = factor(bin, levels = 0:9999))


ggplot(data = t.raster_mhz_sniped, 
       aes(y = Electrode, x = bin)) + 
  geom_raster(aes(fill = spikes), interpolate = F) +
  #geom_vline(xintercept = c(200, 280)) +
  #facet_grid(Treatment + Well ~ ., scales = "free") + 
  facet_wrap(Treatment ~ .,
             scales = "free_y",
             nrow = 2,
             strip.position = "left",
             labeller = as_labeller(c(Scrambled = "Scrambled\nElectrodes",
                                      ShYTFD = "shYthdf123\nElectrodes")))  +
  scale_fill_gradient2(low = "white", high = "black") + 
  scale_x_discrete(breaks = seq(0, 10000, by = 1000), 
                   labels = seq(0, 100, by = 10)) +
  coord_cartesian(xlim = c(0, 9000)) +
  theme(axis.text.y = element_blank(),
        #axis.ticks.y = element_blank(), 
        axis.line.y = element_blank(),
        panel.border = element_rect(size = 0.25, fill = NA)) +
  labs(y = NULL, x = "Recording time, s") +
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        legend.position = "none",
        aspect.ratio = 0.125) -> p.raster_toshow
```
```{r fig.width=6, fig.height=2.5}
ggsave(plot = p.raster_toshow, 
       filename = paste0("./results/", prfx, "Raster.toshow.pdf"),
       width = 6, height = 2.5)
```

## Some bursting and NS statistics
```{r}
t_bursts %>% 
  filter(Well %in% c("D2", "B4"),
         DIV == "18") %>% 
  group_by(Treatment, Well, DIV) %>% 
  summarise(total.bursts = sum(n_bursts),
            avg.brst_length = mean(brst.size_avg),
            nAE = n(), .groups = "drop") %>% 
  mutate(Bursts.p.eltd = total.bursts / nAE)
```
```{r}
t.ns %>% 
  filter(Well %in% c("D2", "B4"),
         DIV == "18")
```




```{r, fig.height=2, fig.width=5}
t.raster_mhz_xtremes <- t.raster_mhz %>% 
  filter(Well %in% c("D1", "C1")) %>% 
  mutate(bin = factor(bin, levels = 0:9999))

ggplot(data = t.raster_mhz_xtremes, 
       aes(y = Electrode, x = bin)) + 
  geom_raster(aes(fill = spikes), interpolate = F) +
  #geom_vline(xintercept = c(200, 280)) +
  #facet_grid(Treatment + Well ~ ., scales = "free") + 
  facet_wrap(Treatment ~ .,
             scales = "free_y",
             nrow = 2,
             strip.position = "left",
             labeller = as_labeller(c(Scrambled = "Scrambled\nElectrodes",
                                      ShYTFD = "shYthdf123\nElectrodes")))  +
  scale_fill_gradient2(low = "white", high = "black") + 
  scale_x_discrete(breaks = seq(0, 10000, by = 1000), 
                   labels = seq(0, 100, by = 10)) +
  coord_cartesian(xlim = c(0, 8000)) +
  theme(axis.text.y = element_blank(),
        #axis.ticks.y = element_blank(), 
        axis.line.y = element_blank(),
        panel.border = element_rect(size = 0.25, fill = NA)) +
  labs(y = NULL, x = "Recording time, s") +
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        legend.position = "none",
        aspect.ratio = 0.125) -> p.raster_extrms
```
```{r fig.width=6, fig.height=2.5}
ggsave(plot = p.raster_extrms, 
       filename = paste0("./results/", prfx, "Raster.extrms.pdf"),
       width = 6, height = 2.5)
```

## Some bursting and NS statistics
```{r}
t_bursts %>% 
  filter(Well %in% c("D1", "C1"),
         DIV == "18") %>% 
  group_by(Treatment, Well, DIV) %>% 
  summarise(total.bursts = sum(n_bursts),
            avg.brst_length = mean(brst.size_avg),
            nAE = n(), .groups = "drop") %>% 
  mutate(Bursts.p.eltd = total.bursts / nAE)
```
```{r}
t.ns %>% 
  filter(Well %in% c("D1", "C1"),
         DIV == "18")
```


# ------------------------------------------------------------------------------
## 4. Network spikes analysis

### Manual filtering (same as before)
```{r}
t.yth_ns <- t_ns %>% 
  filter(!(Well %in% c("D3", "D2") & DIV == "25"),
         Treatment %in% c("Scrambled", "ShYTFD"), 
         !Well %in% c("D5")) %>% 
  mutate(TP = ifelse(DIV %in% c("7", "11", "14"), "DIV <= 14", "DIV >= 18"))

## summarise per DIV for line plotting
t.yth_ns_sum <- t.yth_ns %>%  
  group_by(Treatment, DIV) %>% 
  summarise(NS_br.avg = mean(NS, na.rm = T),
            NS_br.sd = sd(NS, na.rm = T),
            .groups = "drop") %>% 
  mutate(DIV = as.numeric(DIV))
```

```{r}
ggplot(data = t.yth_ns_sum, 
         aes(x = DIV, y = NS_br.avg, color = Treatment)) + 
    geom_point(aes(shape = Treatment), 
               size = 3, 
               position = position_dodge(width = 0.35)) + 
    geom_errorbar(aes(ymax = NS_br.avg + NS_br.sd,
                      ymin = NS_br.avg - NS_br.sd),
                  width = 0.25,
                  position = position_dodge(width = 0.35)) + 
  geom_line(position = position_dodge(width = 0.35)) + 
  #geom_hline(yintercept = 0) +
  scale_color_manual(values = alpha(pal_mut.msa[c(1,3)], 0.67)) + 
  scale_x_continuous(breaks = seq(0, 25, by = 1)) +
  theme(aspect.ratio = 0.5 ,
        legend.position = c(0.2, 0.75)) + 
  #scale_y_log10() +
  #coord_cartesian(ylim = c(0, 1600)) +
  labs(y = "Count of Network Spikes", x = "DIV")
```
```{r}
t.yth_ns %>% 
  mutate(DIV = factor(DIV, levels = c(7, 11, 14, 18, 21, 25))) %>% 
  ggplot(data = ., aes(x = DIV, y = NS)) + 
  geom_boxplot(aes(fill = Treatment)) + 
  scale_y_log10()
```