

```{r Source libraries}
source("./F_process.R")
```

# ------------------------------------------------------------------------------
## Files

Here the script might produce errors if the naming convention in the MEA recording has been changed.
Be vigilant and keep the naming as follows (DIV should be in capitals):
*_DIV21(000)_*.csv
```{r 1. Parse raw data in the source directory }
t.fls <- parse_file.list(".")
```

Assemble a spike list (list)
# ------------------------------------------------------------------------------
```{r Check metadata}
## Here one needs to load any environmental data file
t.meta <- parse_meta("NoBarcode_20221213_Scr_Larp_Elavl_DIV07(000)_environmental_data.csv")
t.meta
```

Here we assemble a main data object from the files that machine spits out.
---- bin_interval parameter can be decreased to increase the resolution (time interval over which spikes are averaged)
default is 0.25, which is 0.25 sec.
---- well_min_rate argument sets a minimum number of electrodes that should be firing for a well to considered actiuve.
Normally, we use 4 electrodes per well, but in some cases the parameter can be reduced to 2.
---- beg = 0 and end = 300 arguments allows to specify the time frame of the recording. 
In our case, the beg is set to 0 and the end is set to 300 (seconds), which corresponds to first 5 mins of recording.

```{r Set up DIVs to iterate over}
## Here one sets up a vector of DIVs to iterate over
divs <- c(7, 11, 14, 18, 21()
```

```{r 2. Assemble table with spikes}
## average spikes
l.div <- list()
for (i in divs) {
  ##
  t.fls %>% 
  filter(data_type == "spike_list", div == i) %>% 
  as.list(.) %>% 
  pmap(.,
       ~ parse_mea(..1) %>%
         filter(Treatment %in% c("Scrambled", "shLarp1", "shElavl234"))) %>% 
    `[[`(1) -> t.spks

  ## Split into a list by well and then by electrode
  t.spks %>% 
    split(., .$Well) %>% 
    purrr::map(., 
               function(welltab) split(welltab, welltab$Electrode) %>% 
                 purrr::map(., 
                            function(eltab) eltab %>% pull(`Time (s)`)
                            )
               ) -> l.wells
  ##
  purrr::map(l.wells,
             ~ construct_s(.x, beg = 0, end = 300, bin_interval = 0.25)) -> l.s
  
  ## Filter 
  purrr::map(l.s,
             ~ filter_spike.list(.x, well_min_rate = 2)) -> l.div[[as.character(i)]]
}

## Drop all empty wells from the list
purrr::map(l.div,
           function(div) {
             purrr::map_lgl(div,
                            ~ length(.x) > 0)
           }) -> l.div_keep
purrr::map2(l.div,
            l.div_keep,
            ~ .x[.y]) -> l.div

## Process list into a table with spike features
purrr::map2_dfr(l.div,
                names(l.div),
                function(el.div, div) {
                  purrr::map2_dfr(el.div, 
                                  names(el.div),
                                  ~ tibble(DIV = div,
                                           Well = .y, 
                                           MFR = mean(.x$meanfiringrate),
                                           stdFR = sd(.x$meanfiringrate),
                                           nae = .x$NCells)
                                  )
                }) %>% 
  inner_join(t.meta, ., by = "Well") -> t.proc
```

## Prepare burst features
# ------------------------------------------------------------------------------
```{r 3. Assemble table with spike bursts}
## read spike burst files
t.fls %>% 
  filter(data_type == "electrode_burst_list") %>% 
  as.list(.) %>% 
  pmap_dfr(.,
           ~ parse_mea(..1) %>%
             filter(Treatment %in% c("Scrambled", "shLarp1", "shElavl234")) %>% 
             mutate(DIV = ..3)) -> t.brst

## assemble a table with Active electrodes
purrr::map2_dfr(l.div,
            names(l.div),
            function(div_l, div) {
              purrr::map2_dfr(div_l,
                          names(div_l),
                          function(well_tab, well) {
                            tibble(DIV = div,
                                   Well = well, 
                                   AE = well_tab$channels)
                          })
            }) -> t.ae_keep
## Filter burst table by active electrodes and restrict the time frame to 5 minuts (before stimulation)
t.brst %>% 
  mutate(DIV = as.character(DIV)) %>% 
  inner_join(t.ae_keep,
             ., 
             by = c("DIV", "Well", "AE" = "Electrode")) %>% 
  filter(`Time (s)` <= 300) %>%   ## <-----------------------------------
  group_by(Well, Treatment, DIV, AE) %>% 
  summarise(n_bursts = n(),
            brst.size_avg = mean(`Size (spikes)`),
            brst.length_avg = mean(`Duration (s)`),
            .groups = "drop") -> t.brst_eltd
```

## Prepare Network spikes (NS) features
# ------------------------------------------------------------------------------
```{r 4. Assemble table with network spikes}
purrr::map2_dfr(l.div,
                names(l.div),
                function(div_l, div) {
                  purrr::map2_dfr(div_l,
                              names(div_l),
                              function(well_tab, well) {
                                n.spks <- well_tab$firing_rates$frate
                                n.ns <- sum(n.spks >= 4)
                                tibble(DIV = div,
                                       Well = well, 
                                       NAE = well_tab$NCells,
                                       NS = n.ns)
                              })
                }) -> t.ns
## add meta data
t.ns <- t.ns %>% 
  inner_join(t.meta, ., by = "Well") %>% 
  mutate(DIV = factor(DIV, levels = divs))
```


## Write down results
# ------------------------------------------------------------------------------
```{r 5. Write down processed results}
if (!dir.exists("./processed")) {
  dir.create("./processed")
}

## list of all results
write_rds(l.div, file = "./processed/l_assembled.RDS")

## Table with well annotation
write_tsv(t.meta, file = "./processed/t_meta.tsv")
## Spike table
write_tsv(t.proc, file = "./processed/t_spikes.tsv")
## Burst table
write_tsv(t.brst_eltd, file = "./processed/t_bursts.tsv")
## NS table
write_tsv(t.ns, file = "./processed/t_ns.tsv")
```


